---
title: "Quiz 9 R Markdown"
author: "Bella Colpo"
date: "5/25/2018"
output: html_document
---

```{r, echo = FALSE, include = FALSE}
#Install dplyr and ggplot2
#install.packages("dplyr")
library(dplyr)
#install.packages("ggplot2")
library(ggplot2)
#Read Puget Sound Basin Data and rename "psbasin"
#
```


#**Plot 1: pH and Depth**#


```{r, echo = FALSE}
psbasin <- read.csv("Data/PugetSoundBasin_WP_July16.csv")



#
#Filter the psbasin data to just include the July datasets
july_psbasin <- filter(psbasin,date =="7/11/16")

#Select pertinent columns for water quality
july_psbasin <- select(july_psbasin, depth..meters., salinity..psu., temperature..centigrade., chlorophyll.raw..ug.l., dissolved.oxygen..mg.l..raw, pH)
#
#Print modified data table with depth, salinity, temperature, chlorophyll, dissolved oxygen, and pH
#print(july_psbasin)
#
#Rename and read Port Townsend data to reflect the same columns as the july_psbasin data
pt <- read.csv("Data/PortTownsend_16.csv")

july_pt <- filter(pt, date == "7/6/16")
july_pt <- select(july_pt, depth..meters., salinity..psu., temperature..centigrade., chlorophyll.raw..ug.l., dissolved.oxygen..mg.l..raw, pH)

#Repeat same process as I dud with the Port Townsend and Puget Sound Basin datasets with the third dataset: Elliot Bay

eb <- read.csv("Data/ElliotBay_16.csv")


july_eb <- filter(eb, date == "7/11/16")
july_eb <- select(july_eb, depth..meters., salinity..psu., temperature..centigrade., chlorophyll.raw..ug.l., dissolved.oxygen..mg.l..raw, pH)

#Combined the three datasets into one large dataset using the bind_rows function from dplyr, and renamed the three datasets to specify the locations and used .id to create a new location column.
full_data <- bind_rows(list(psbasin = july_psbasin, eb = july_eb, pt = july_pt), .id = "location")
#Plot the pH for all three locations by depth using ggplot
ggplot(full_data, aes(x = pH, y = depth..meters., color = location)) + geom_point(data = full_data, aes(x = pH, y = depth..meters.)) + labs (x = "pH", y = "Depth (m)")
```

The first plot I created represents the data from [Washington State Department of Ecology Water Quality Data](https://fortress.wa.gov/ecy/eap/marinewq/mwdataset.asp?theyear=2014&themonth=&ec=no&scrolly=21&htmlcsvpref=csv&estuarycode=1&staID=121). Using this dataset from the Washington Department of Ecology, I wanted to see how water quality during the summer of 2016. This was at the tail-end of a few summers of harmful algal bloom outbreaks, so I want to see how water quality was different in a few areas of the Puget Sound. I did not separate the locations in the same graph here, but will do so in the next plot.


#**Plot 2: pH and Depth with facet_wrap**#


```{r, echo = FALSE}
#ggsave("pH_by_depth.jpg")
#Tried another way usig the facet_wrap. 
ggplot(full_data) +  geom_point(aes(x = pH, y = depth..meters.)) + facet_wrap(~location) + labs(x = "pH", y = "Depth (m)") + ggtitle("Puget Sound pH in Elliot Bay, Puget Sound Basin, and Port Townsend")
#ggsave ("version2_nocolor_pH_and_depth.jpg")
```

This plot above represents the same data that I isolated from the 2016 for the three locations: Elliot Bay, Puget Sound Basin, and Port Townsend.  The plot above, however, uses facet_wrap to separate the three locations in visualization.


#**Plot 3: Improved Plots with Reversed X and Y Axis**#


```{r, echo = FALSE}
#Exercise 1: Generating improved plots
#First Plot: Switching x and y axis so plot is more readable


ggplot(full_data) +  geom_point(aes(x = depth..meters., y = pH)) + facet_wrap(~location) + labs(x = "Depth (m)", y = "pH") + ggtitle("Puget Sound pH in Elliot Bay, Puget Sound Basin, and Port Townsend")
#ggsave ("depthandpH.jpg")
```

This plot above shows the x and y axis switched from the previous two, representin the pH and depth of the three sites that were surveyed.


#**Plot 4: Averages of Each Location**#



```{r, echo = FALSE}
#Second Plots:

# Turn averages of each location into new data frame
aves <- full_data %>% 
  group_by(location) %>% 
  summarise(pH = mean(pH))

# Plot as bar plot as suggested by a peer last week
ggplot(data = full_data, aes(x = location, y= pH)) + 
  coord_cartesian(ylim=c(7.5,8)) + 
  geom_bar(data = aves, stat = "identity") + labs(x = "Location", y = "pH")
```